### 2015-07-07 narrative/data-reduction.tex

\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owing to the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
docker container so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. A priori information on the nature of the data is
used to choose the basic algorithm while the actual values of the data are used
to tune the parameters used in the reconstruction procedures. We shall refer to
this paradigm as an Auditor.  We will examine the use of several auditors for
the simulations which we will work with. One approach is to use a fluid-code to
audit a kinetic code, which can be a good approximation over short periods of
time. Once there is a large deviation (specified by the user) from the auditor
to the original calculation, the data will be saved. This is a local operation,
since the deviations will grow differently in space and in time.

%%%%%%%%%%%%%
We will be looking at a multi-throng
approach to data-refactoring. One approach allows the user to specify
information about their data to indicate if the data is spatially varying,
space-and time varying, just time varying. This allows us to store information
in NVRAM on each node (when and where necessary) to later preform compression
in both space and time.

>>>>

\paragraph{High Impact Challenges:} Maximal data reduction on the SSIO layer
would result from storing only the code and the input parameters, and simply
recalculating all of the results. This extreme is not feasible in practice
owing to the size of the computation involved, but sheds light on the path to
achieving a step change in data reduction. We shall investigate the possibility
of writing a minimal set of information, and placing an ``auditor'' code in a
docker container so that data can be re-generated with a {\em local}
approximation of the full simulation. This is an entirely new approach to
tackle the data overload by using preprocessors in which information is not
simply stored in the form of raw data, but rather in the form of appropriate
parameterized algorithms. A priori information on the nature of the data is
used to choose the basic algorithm while the actual values of the data are used
to tune the parameters used in the reconstruction procedures. We shall refer to
this paradigm as an Auditor.  We will examine the use of several auditors for
the simulations which we will work with. One approach is to use a fluid-code to
audit a kinetic code, which can be a good approximation over short periods of
time.

### 2015-07-07 narrative/deliverables.tex
getting changes for the evaluatoin

\section*{Deliverables}

\begin{itemize}
\item plug-in architecture for Sirocco
\item storage APIs to manipulate storing and retrieving data to specific tiers
\item metadata management for data stored in multiple tiers using different compression
\item metadata management for deeper data knowledge
\item middleware hooks to manage new storage APIs
\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}

>>>>

\section{Deliverables}
The following are the per year project deliverables:

\textbf{Year 1}
\begin{itemize}
\item Develop new techniques for data description that allow users to 
describe the data utility based on their expectations. (ORNL)

\item Assemble and evaluate general techniques for refactoring of data
that will be embedded in the system and can be selected by users or applications. (Brown/ORNL)
\item Explore the use of application hints and utility of data to guide the initial placement of data. 
(Rutgers/ORNL)
\item Investigate the trade-off between ``filing'' and ``piling'' data. (UCSC/Sandia)
\item Demonstrate a metadata service capable of serving both POSIX clients and our clients,
but without consideration for authentication, authorization, or data migration. (Sandia/UCSC)
\end{itemize}

\textbf{Year 2}
\begin{itemize}
\item Develop a new type of querying system that allows an application to ask the storage system
about completion timing information. (ORNL)
\item Research autonomic data management strategies that
can evaluate utility/cost tradeoff, and appropriately place/move data objects at runtime.
\item Demonstrate a time bounded search approach for finding data within the storage system
to identify data and the current location. This will update the cached information in the metadata
service to short-circuit future requests for the data presuming that it does not migrate again soon. 
\item Extend Sirocco to enforce data-centric metrics that help guide the QoS decisions.
(Sandia/UCSC)
\end{itemize}

\textbf{Year 3}
\begin{itemize}
\item Work with DOE applicationexsiting ADIOS-enabled applications, framework using many of the applications to which we have access, including XGC1, GTC, and SPECFM3D
%\item storage APIs to manipulate storing and retrieving data to specific tiers
%\item metadata management for data stored in multiple tiers using different compression
%\item metadata management for deeper data knowledge
%\item middleware hooks to manage new storage APIs
%\item example lossy and lossless compression plug-ins to demonstrate the effectiveness of this approach
\end{itemize}

### 2015-07-07 narrative/enabling_tech.tex

We will also take advantage of the many innovations discovered in the Ceph~\cite{weil:osdi06} storage
system project. 
%
%  In the LWFS project, Sandia sought to demonstrate stripping
%down a file system to the bare minimum features required for any storage
%system. Then, additional functionality and semantics, such as POSIX-style
%directories and consistency requirements, were added in using auxiliary
%services separate from the stack or by layering providing a new interface to
%the storage system. The project's first phase demonstrated that using a core
%object storage system with integrated security providing authentication and
%authorization services, customized storage interfaces could be developed with
%exactly the overhead required for different applications. The second phase of
%LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
%devices for efficient and scalable storage performance.  Sirocco has been in
%process for three years and has developed a usable storage infrastructure.
%
The Ceph project (ceph.com) started at UC Santa Cruz in 2004 and was initially
funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the goal to create an
object-based parallel file system that addresses the well-known metadata
service bottleneck of creating and accessing 1000s of files from a typical 
checkpoint~\cite{weil:osdi06}.  However, after the Ceph kernel client was
pulled into the Linux mainline in 2010, the focus of Ceph shifted to its
object-based storage subsystem, also called Reliable Autonomous Distributed
Object Store (RADOS), which now is the dominant storage system in OpenStack
production-level deployments and was recently selected as the new backend for
Flickr and Yahoo! Cloud Services. In this project we will leverage two of the
key innovations of Ceph: (1) Controlled Replication Under Scalable Hashing
(CRUSH)~\cite{weil:sc06} will help make resource identification and
predictable performance scalable, and (2) extensible storage object
classes~\cite{watkins:ucsctr15} will provide the main mechanisms of introducing
new APIs, metadata vocabulary, and quality of service into the storage stack.

>>>>

We will also take advantage of the many innovations discovered in the Ceph~\cite{weil:osdi06} storage
system project. 
%
%  In the LWFS project, Sandia sought to demonstrate stripping
%down a file system to the bare minimum features required for any storage
%system. Then, additional functionality and semantics, such as POSIX-style
%directories and consistency requirements, were added in using auxiliary
%services separate from the stack or by layering providing a new interface to
%the storage system. The project's first phase demonstrated that using a core
%object storage system with integrated security providing authentication and
%authorization services, customized storage interfaces could be developed with
%exactly the overhead required for different applications. The second phase of
%LWFS, known as Sirocco~\cite{sirocco}, is investigating how to manage storage
%devices for efficient and scalable storage performance.  Sirocco has been in
%process for three years and has developed a usable storage infrastructure.
%
The {\bf Ceph} project (ceph.com) started at UC Santa Cruz in 2004 and was initially
funded by DOE/NNSA involving LLNL, LANL, and Sandia, with the goal to create an
object-based parallel file system that addresses the well-known metadata
service bottleneck of creating and accessing 1000s of files from a typical 
checkpoint~\cite{weil:osdi06}.  However, after the Ceph kernel client was
pulled into the Linux mainline in 2010, the focus of Ceph shifted to its
object-based storage subsystem, also called Reliable Autonomous Distributed
Object Store (RADOS), which now is the dominant storage system in OpenStack
production-level deployments and was recently selected as the new backend for
Flickr and Yahoo! Cloud Services. In this project we will leverage two of the
key innovations of Ceph: (1) Controlled Replication Under Scalable Hashing
(CRUSH)~\cite{weil:sc06} will help make resource identification and
predictable performance scalable, and (2) extensible storage object
classes~\cite{watkins:ucsctr15} will provide the main mechanisms of introducing
new APIs, metadata vocabulary, and quality of service into the storage stack.
